\LARGE \bf On Physics-Based 3D Waypoint Generation for Autonomous Exploration in Mobile Robotics

### 2011-08-30 paper/On Physics-Based 3D Waypoint Generation/paper.tex
zeugs
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1097 f4953187-ecf7-0310-90f3-8d0fedb0aae5

This paper presents a novel approach on increasing the efficiency of such mapping procedures. While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view. Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces. This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints. The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of about 15 minutes. Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.

@@

Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments []. Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.

In general, frontier-based approaches are a popular method to computing new waypoints [1,2,3,4], as near-frontier sensor-poses often promise high information gains. Unfortunately, information about frontiers between known and unknown space is hard to generate from three-dimensional datastructures such as pointclouds or octrees.

Occupancy grid maps are often used bound to 2d environments or voxel structures.

The next-best-view problem is similar, but does not include the constraints of safe navigation. Extension to art-gallery-problem.

Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form. Although the data structure permits easy traversability computation, storing more complex geometry or overhangs remains impossible, limiting their practical use to two-dimensional environments. Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption. Multi-Level surface maps, introduced by Triebel et al [triebel], eliminate many of these limitations, while still being constrained to a grid.

e.g. ``Autonomous Exploration for 3D Map Learning'' by Johoo et al.

@@

As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data. Finding the next best view in such a point cloud can be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals. Instead of trying to generate these details from the point cloud whenever it changes, our algorithm does not require such information.

@@

Although any geometry can be used for waypoint generation, using spheres yields three important advantages:
\begin{itemize}
  \item Spheres can represented only by radius and position, so the overall memory requirement is very low (especially when all spheres share a common radius).
  \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the narrowphase in collision detection. To detect a collision, it is sufficient to check whether the distance between the sphere's center and the point is smaller than the sphere's radius.
  \item Most importantly, spheres are ideally suited to slip through smaller ``holes'' in the pointcloud, enabling their detection.
\end{itemize}

Any sensor generating spatial occupancy information can be used, as the algorithm's input consists solely of geometry.

Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.

Simulated annealing.

\subsection{3D Model / Data structures}

point cloud, octree

@@

@@

With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection. In general, physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

The authors gratefully acknowledge the contribution of TAMS.

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

@@

>>>>

This paper presents a novel approach on increasing the efficiency of such mapping procedures. While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view. Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces[TODO]. This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints. The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of about 15 minutes. Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.

@@

Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [...]. Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.

Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem, as it does not include the constraints of safe navigation. In general, frontier-based approaches are a popular method to compute new waypoints [1,2,3,4] as sensor-poses from areas between known and unknown environment offer a good compromise between safe reachability and high information gain. Unfortunately, information about frontiers between known and unknown space is hard to generate from three-dimensional datastructures such as pointclouds or octrees.

\subsection{Data structures}

Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form. Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains impossible, limiting its practical use to two-dimensional environments. Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption. Multi-Level surface maps, introduced by Triebel et al [triebel], eliminate many of these limitations, while still being constrained to a grid.

@@

As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data. Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input. Finding the next best view in such data be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals. Instead of trying to generate these details from the point cloud whenever it changes, our algorithm does not require such information.

@@

Although any geometry can be used for waypoint generation, using spheres yields three important advantages:
\begin{itemize}
  \item Spheres can represented only by radius and position, so the overall memory requirement is very low (especially when all spheres share a common radius).
  \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the midphase in collision detection. To detect a collision, it is sufficient to check whether the distance between the sphere's center and the point is smaller than the sphere's radius.
  \item Most importantly, spheres are ideally suited to slip through smaller ``holes'' in the pointcloud, enabling their detection.
\end{itemize}

%Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.

%Simulated annealing.

@@

The algorithms complexity derives from the computational effort of the collision detection phase in physics simulation and thus from the number of collision objects. Collision detection between $n$ objects in general requires $n(n-1)/2$ collision checks to be performed, leading to a complexity of O($n^2$). Optimized algorithms like sort-and-sweep or implementations relying on spatial subdivison may reduce the complexity to O($n$ log $n$) in favorable cases \cite{legrand2007}. This initially appears to be a major obstacle to employment of the algorithm outside of simulation, as the pointcloud may grow to several million points during a scan. Fortunately, the following optimizations have made the algorithm sufficiently fast for real-time application:

\begin{itemize}
  \item Given a pointcloud consisting of $n$ points and a set of $m$ sample geometries, the algorithm does not require collision tests between all $n+m$ objects. As the $n$ points making up the accumulated scan data are static, the number of required tests is reduced to $\frac{m(m-1)}{2} + n*m$. Because $n >> m$, this optimization yields a considerable loss in computational effort.
  \item When using the sensordata solely for waypoint generation, the density of the pointcloud can be reduced to allow gaps almost the size of the sample geometry's radius. That is, if sample spheres are created with a radius of e.g. 1m, the interacting pointcloud may discard points if there are neighbors within a distance of less than ~2m. In practice, this means a reduction of $n$ by two or three dimensions. For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the needed threshold.
  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculationi on the CPU.
\end{itemize}

@@

With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection. Physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs

@@

%\section{ACKNOWLEDGMENTS}
%The authors gratefully acknowledge the contribution of TAMS.

@@

% ``Autonomous Exploration for 3D Map Learning'' by Johoo et al.

\bibitem{legrand2007}
% macht broadphase per CUDA, gut erklärt in GPU gems 3, chapter 32
Scott Le Grand, Broad-Phase Collision Detection with CUDA, {\it GPU Gems 3, } (2007)
